### 2016-02-02 paper/main.tex
fix minor typos from David Poznick

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
simply estimation with some positive prior probability that the effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
the $\lfsr$ and $\lfdr$ can be quite different, and emphasize benefits of the $\lfsr$, particularly its increased robustness to modeling assumptions.

@@

In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can be used, and we found this to perform adequately for moderate-sized problems, say $J<2000$). And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

@@

The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
\begin{equation} \label{eqn:lfsr}
\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
To illustrate, suppose that 
\begin{gather*}
p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
\end{gather*} 
Then 
from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.

@@

\begin{figure}
\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

@@

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0. See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
demonstration.

@@

The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.

@@

An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
can estimate the {\it size} of each effect $\beta_j$.
Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
threshold $T$.
%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
%towards zero, with the strength of the shrinkage being estimated from the data.
Further, because the posterior distribution is, by definition,
conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
how can confidence intervals be assigned to the true
effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
%but also notes the difficulties of implementing a stable general algorithm.

>>>>

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
simply estimation with some positive prior probability that the effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
``local false sign rate" (\lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
the $\lfsr$ and $\lfdr$ can be quite different, and emphasize benefits of the $\lfsr$, particularly its increased robustness to modeling assumptions.

@@

In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  \left[\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2) \right],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can be used, and we found this to perform adequately for moderate-sized problems, say $J<2000$). And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

@@

The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey \cite{tukey1991philosophy} argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey (\cite{tukey1962future}, p32) suggested  that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
\begin{equation} \label{eqn:lfsr}
\lfsr_j := \min[ p(\beta_j \geq 0| \hat\pi, \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
To illustrate, suppose that 
\begin{gather*}
p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
\end{gather*} 
Then 
from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.

@@

\begin{figure}
\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the ways different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

@@

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (cyan). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the cyan histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the ZA pushes the dark blue component until the cyan component {\it disappears} at 0. See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
demonstration.

@@

The key point we want to make here is that estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.

@@

An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
can estimate the {\it size} of each effect $\beta_j$.
Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
threshold $T$.
%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
%towards zero, with the strength of the shrinkage being estimated from the data.
Further, because the posterior distribution is, by definition,
conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
how can confidence intervals be assigned to the true
effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
and \cite{zhao2012empirical} considers in detail the case where the non-null effects are normally distributed.
%but also notes the difficulties of implementing a stable general algorithm.
